---
title: "Tutorial"
output: 
  learnr::tutorial:
    progressive: true # reveals one content section at a time.
    allow_skip: true # students can skip sections or move on without completing a section
    toc: true
    toc_depth: 2
    toc_float: yes
    theme: journal
runtime: shiny_prerendered
---

```{r setup, include=FALSE, exercise.checker=TRUE }

library(learnr)

library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
EatTweets <- read_csv("C:/Users/aleaw/OneDrive/Documents/GitHub/PPOL105/EatTweets.csv")
```

This data is from the .....
```{r}
EatTweets # as a tibble
```
```{r arrange, exercise = TRUE, eval=TRUE}
# change the sort order to descending
arrange(EatTweets, Year)
```
```{r Filter, exercise = TRUE, eval=TRUE}
# change the filter to only at 1990. Then try 2012.
filter(EatTweets, Year == "1985")
```

```{r Select, exercise  = TRUE}
# change the code to only look at variables for the USA
select(EatTweets, contains ("UK"))
```
```{r}
custom_checker <- function(label, user_code, check_code, envir_result, evaluate_result, last_value, ...) {
  # this is a code check
  if (is.null(envir_result)) {
    if (is_bad_code(user_code, check_code)) {
      return(list(message = "I wasn't expecting that code", correct = FALSE))
    }
    return(list(message = "Nice code!", correct = TRUE))
  }
  # this is a fully evaluated chunk check
  if (is_bad_result(last_value, check_code)) {
    return(list(message = "I wasn't expecting that result", correct = FALSE))
  }
  list(message = "Great job!", correct = TRUE, location = "append")
}

tutorial_options(exercise.checker = custom_checker)
```


## Wealth Change Example {#wealthchange}
```{r Plot}  
wealthchange <- EatTweets %>% ggplot(aes(x = Year, y = USA_Pretax_income_top1)) + geom_point()

wealthchange + annotate( "text", x = 1975, y=.2, label = "Average wealth for top 1%") # adds text at that location on the graph

wealthchange + annotate("rect", ymin = .13, ymax = .15, xmin = 1980, xmax=1990, alpha = .4) # makes a rectangle on the graph

```

Bookdown Tips:
https://happygitwithr.com/bookdown-cheat-sheet.html#cross-references
```{r}
# to refer to a different chunk or location in the text
#\@ref(wealthchange) # in bookdown

```

## Topic 1

### Exercise 

*Here's a simple exercise with an empty code chunk provided for entering the answer.*

Write the R code required to add two plus two:

```{r two-plus-two, exercise=TRUE}

```

### Exercise with Code

*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaulated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
EatTweets
```

```{r print-limit-hint}
head(EatTweets)
```

## Quizes!

```{r quiz}
# *You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

quiz(
  question("What's Alea's favorite animal?",
    answer("Penguins"),
    answer("Lions",  correct = TRUE),
    answer("Aardvarks"),
    answer("Groundhogs")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

