---
title: "Crashes & COVID19"
author: "Alea Wilbur"
date: "3/1/2021"
output: html_document
#runtime: shiny
---

# Setup
```{r setup, include=FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(lubridate) 
library(plotly)
library(sf)
library(leaflet) 
library(tmap)
library(highcharter) 
library(ggsci)
library(kableExtra) # make your charts pretty
#devtools::install_github("covid19R/covid19nytimes")
library(covid19nytimes)
library("RSocrata")


knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
#crash <- read_csv("~/GitHub/PPOL105/Traffic_Crashes_Cleaned.csv")
#crashes <- read_csv("C:/Users/aleaw/OneDrive/Desktop/LessonPlans/Original Datasets_Dont Touch/Traffic_Crashes_Orig.csv")

```

```{r}

df <- read.socrata(
  "https://data.cityofchicago.org/resource/85ca-t3if.csv", 
  app_token = "NLJIazfKKzk71gTZXgyJ17H8N",
  email     = "awilbu2@uic.edu",
  password  = "Supersoup!55"
)

df2 <- read.socrata(
  "https://data.cityofchicago.org/resource/85ca-t3if.csv?$where = crash_date>2019-01-01T00:00:01", 
  app_token = "NLJIazfKKzk71gTZXgyJ17H8N",
  email     = "awilbu2@uic.edu",
  password  = "Supersoup!55"
)
#API KEY ID
# 7wozsduhlxqle390n4kjkd7a5

# API Key Secret
# 49p2nh9108wuuine4z720roen4x6yyd2wd5wtn8qdennfrsksb
```

This [link](https://rpubs.com/camilovelezr/tcc) has done most of the work for crashes

[Example with utah and religions] (https://statsf18.classes.andrewheiss.com/files/your-name_problem-set-2.Rmd) 

```{r}
crash <- crashes 
#%>%  # READ IN ORIGINAL FILE 
#select(-c(CRASH_RECORD_ID, CRASH_DATE_EST_I, TRAFFIC_CONTROL_DEVICE, DEVICE_CONDITION, LANE_CNT, ALIGNMENT, ROAD_DEFECT, REPORT_TYPE, INTERSECTION_RELATED_I:HIT_AND_RUN_I, DATE_POLICE_NOTIFIED, PHOTOS_TAKEN_I:DOORING_I,NUM_UNITS, INJURIES_INCAPACITATING:INJURIES_REPORTED_NOT_EVIDENT, INJURIES_UNKNOWN)) 

crash$CRASH_DATE <- mdy_hms(crash$CRASH_DATE)  # tells R that date is in month/day/year and hour:minute:second formatting

crash$CRASH_DATE <- as_date(crash$CRASH_DATE) # Remove hours, I did this in excel already for the cleaned version. 
# Time also exists already in another column
# Date is in year/month/day format.
```


```{r}

min(crash$CRASH_DATE) # March 3rd, 2013

crash <- crash %>%
  filter(CRASH_DATE >= (mdy("1/1/2019"))) # 222,717 observations remaining

min(crash$CRASH_DATE) # January 1, 2019
```
So far, so good!

```{r}
# data %>%
#   mutate(educcat_F = fct_recode(educcat_F, #<<
#                                 "1" = "Less Than High School", #<<
#                                 "2" = "High School", #<<
#                                 "3" = "Junior College",#<<
#                                 "4" = "Bachelor", #<<
#                                 "5" = "Graduate")) %>% #<<
#   group_by(educca t_F) %>%
#   summarize(count = n())
# class(crash$POSTED_SPEED_LIMIT)
# crash %>% mutate(POSTED_SPEED_LIMIT == recode(POSTED_SPEED_LIMIT, 
#                                               1:7 = 5,
#                                               8:12 = 10)) 
#                  POSTED_SPEED_LIMIT == 13:17, 15, POSTED_SPEED_LIMIT | 
#                  POSTED_SPEED_LIMIT == 18:22, 20, POSTED_SPEED_LIMIT) | 
#                  POSTED_SPEED_LIMIT == 23:27 -> 25,
#                  POSTED_SPEED_LIMIT == 28:32 -> 30,
#                  POSTED_SPEED_LIMIT == 33:37 -> 35,
#                  POSTED_SPEED_LIMIT == 38:42 -> 40,
#                  POSTED_SPEED_LIMIT == 43:47 -> 45,
#                  POSTED_SPEED_LIMIT == 48:52 -> 50,
#                  POSTED_SPEED_LIMIT == 53:57 -> 55,
#                  POSTED_SPEED_LIMIT == 58:62 -> 60,
#                  POSTED_SPEED_LIMIT == 63:70 -> 65)
```

```{r}
#crashes per month
count <- crash %>% 
  group_by(year= year(CRASH_DATE), month= month(CRASH_DATE)) %>% 
  summarize(n=n())

count %>% 
  ggplot(aes(x = year, y = n)) +
  geom_bar(stat = "identity") +
  scale_x_continuous(name = "Number of Crashes Per Year")

```

```{r}
count$date <- mdy(paste0(count$month,"-1-",count$year)) #Date Column
count <- count %>% 
  rename(Date=date, Count=n, Month=month, Year=year)
hc <- highchart(type="stock") %>%
  hc_add_series(data = count,
                   type = "line", hcaes(x = Date,
                   y = Count, 
                   )) %>% 
    hc_xAxis(title = list(text="Date")) %>% 
  hc_colors("#007fff") %>% 
  hc_plotOptions(series=list(backgroundColor = "#535353")) %>% 
    hc_tooltip(borderColor = "black",
             pointFormat = "<b> Crashes</b>: {point.Count}<br>")
hc
```

Speed limits were not recorded in common increments of 5pmh. 
```{r}
# original crash speed limit data
crash %>% 
  group_by(speed_limit=POSTED_SPEED_LIMIT) %>% 
  summarize(n=n()) %>%
  kbl(caption = "Crashes Occuring In Each Speed Limit within Chicago") %>% 
  kable_classic( full_width = T) %>%
  column_spec(1,  # column 1
              bold=T,  # make it bold
              border_right = T) %>% # add the line on the right side of the column
  column_spec(2)
```

Now include only increments of 5mph:
```{r}
slc <- crash %>% 
  filter(POSTED_SPEED_LIMIT %in% seq(from=5, to=70, by=5)) %>% 
    group_by(speed_limit=POSTED_SPEED_LIMIT) %>% 
    summarize(n=n())

kbl(slc) %>%
  kable_classic(full_width = F) %>%
  add_header_above(c("MPH" = 1, "Crashes" = 1)) # number of items in each group and gives them a label

# cooler if you are trying to group multiple variables in a chart.
```
Really good link for Kable charts:
https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html

## Crashes within each posted speed limit
```{r}
n1 <- slc %>% 
  ggplot(aes(x = speed_limit, y = n)) +
  geom_bar(stat = "identity", fill="#68ccff")+
  scale_x_continuous(breaks = seq(5,70,5), labels = seq(5,70,5)) +
  scale_y_continuous(labels = scales::comma) + 
  geom_text(aes(label = n), vjust = -0.3) + 
  ylab("Crashes") +
  xlab("Posted Speed Limit (mph)") +
  theme(panel.background = element_rect(fill = "#e8f7ff"), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
n1 # get rid of decimal point later
```

Bring in the [Chicago Beats Shapefile](https://data.cityofchicago.org/w/aerh-rz74/3q3f-6823?cur=wfOoWMuz9-v&from=root) for mapping!
```{r}
shapefile <- st_read("C:/Users/aleaw/OneDrive/Desktop/LessonPlans/LessonPlans_PPOL105/crashes-COVID19_files/geo_export_19c4ac07-2232-427b-9ba0-127c6c22cd08.shp")

```


# Map
```{r}
beatcount <- crash %>% 
  group_by(BEAT_OF_OCCURRENCE) %>% 
  summarize(count=n()) %>% 
  arrange(BEAT_OF_OCCURRENCE) # Creating an object  with the count of crashes in each beat

# practice map
tm_shape(shapefile) +
  tm_borders() +
  tm_shape(shapefile) +
  tm_fill(col = "blue", alpha=0.65)

shapefile$beat_num <- as.numeric(shapefile$beat_num)

#beatmap <- merge(shapefile, beatcount )
beatmap <- inner_join(shapefile, beatcount, by= c("beat_num" = "BEAT_OF_OCCURRENCE"))

beatmap %>% rename(Crashes=count) %>% qtm("Crashes")
```

# Crash Statistics

```{r}
table(crash$WEATHER_CONDITION)

#crash %>% summarize(WEATHER_CONDITION) %>% kbl()
```

```{r}
table(crash$LIGHTING_CONDITION)
```

```{r}
crash$WEEKDAY <- weekdays(crash$CRASH_DATE) %>% as_factor()
table(crash$WEEKDAY)
```


# Leaflet Map
```{r}
beatmap <- st_transform(beatmap, "+proj=longlat +datum=WGS84") %>% st_zm()

redtheme <- colorNumeric(palette = "Reds", domain=beatmap$count)

leaflet(beatmap) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(stroke = F, 
              smoothFactor = 0.2,
              fillOpacity = .8, 
              color = ~redtheme(beatmap$count))
```


```{r}
table(crash$WEEKDAY, crash$WEATHER_CONDITION)
```

# COVID-19 Data

[Info](https://github.com/Covid19R/covid19nytimes)

```{r}
head(covid19nytimes_states_demo) %>% knitr::kable()
head(covid19nytimes_counties_demo) %>% kable()
```


```{r}
states_covid <- refresh_covid19nytimes_states() # 48,198 observations, 7 variables


# if link is broken load from github folder
# load("../data/us_states_long.rdata")
territories <- c("Guam", "Northern Mariana Islands")
us_states_long <- states_covid %>% filter(!(location %in% territories))

# save(us_states_long, file = "us_states_long.rdata")
us_states_long %>%
  head() %>%
  kable()
```


```{r}
counties_covid <- refresh_covid19nytimes_counties() # BIG file, 2,697,942 obs, 7 variables
```

```{r}
covid19nytimes_states %>%
  filter(location %in% c("Washington", "New York", "Massachusetts", "Michigan", "Illinois")) %>%
  filter(data_type == "deaths_total") %>%
  ggplot(aes(x = date, y = value, color = location)) +
  geom_line() +
  theme_minimal(base_size=14) +
  scale_y_continuous()
```


# Sources
[Chicago Traffic Crash Data](https://data.cityofchicago.org/Transportation/Traffic-Crashes-Crashes/85ca-t3if)

[COVID-19 Data](https://github.com/Covid19R/covid19nytimes)


```{r}

```


# Example HW from online

- Who did you collaborate with: xxxx
- Approximately how much time did you spend on this problem set: xxxx
- What, if anything, gave you the most trouble: xxxx


# Load and wrangle data

Load two CSV files and join them to create a data frame named `crashes` (i.e. using `crashes` as the main data frame, left join `COVID19` to it)

```{r load-libraries-data}

#crashes <- read_csv("data/  .csv")
#COVID19 <- read_csv("data   .csv")
```
