---
title: "Crashes & COVID19"
author: "Alea Wilbur"
date: "3/1/2021"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(lubridate) 
library(plotly)
library(sf)
library(leaflet) 
library(tmap)
library(highcharter) 
library(ggsci)
library(kableExtra) # make your charts pretty

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
#crash <- read_csv("~/GitHub/PPOL105/Traffic_Crashes_Cleaned.csv")
crashes <- read_csv("C:/Users/aleaw/OneDrive/Desktop/LessonPlans/Original Datasets_Dont Touch/Traffic_Crashes_Orig.csv")

```


This [link](https://rpubs.com/camilovelezr/tcc) has done most of the work for crashes

[Example with utah and religions] (https://statsf18.classes.andrewheiss.com/files/your-name_problem-set-2.Rmd) 

```{r}
crash <- crashes 
#%>%  # READ IN ORIGINAL FILE 
#select(-c(CRASH_RECORD_ID, CRASH_DATE_EST_I, TRAFFIC_CONTROL_DEVICE, DEVICE_CONDITION, LANE_CNT, ALIGNMENT, ROAD_DEFECT, REPORT_TYPE, INTERSECTION_RELATED_I:HIT_AND_RUN_I, DATE_POLICE_NOTIFIED, PHOTOS_TAKEN_I:DOORING_I,NUM_UNITS, INJURIES_INCAPACITATING:INJURIES_REPORTED_NOT_EVIDENT, INJURIES_UNKNOWN)) 

crash$CRASH_DATE <- mdy_hms(crash$CRASH_DATE)  # tells R that date is in month/day/year and hour:minute:second formatting

crash$CRASH_DATE <- as_date(crash$CRASH_DATE) # Remove hours, I did this in excel already for the cleaned version. 
# Time also exists already in another column
# Date is in year/month/day format.
```


```{r}

min(crash$CRASH_DATE) # March 3rd, 2013

crash <- crash %>%
  filter(CRASH_DATE >= (mdy("1/1/2019"))) # 222,717 observations remaining

min(crash$CRASH_DATE) # January 1, 2019
```
So far, so good!

```{r}
# data %>%
#   mutate(educcat_F = fct_recode(educcat_F, #<<
#                                 "1" = "Less Than High School", #<<
#                                 "2" = "High School", #<<
#                                 "3" = "Junior College",#<<
#                                 "4" = "Bachelor", #<<
#                                 "5" = "Graduate")) %>% #<<
#   group_by(educcat_F) %>%
#   summarize(count = n())
# class(crash$POSTED_SPEED_LIMIT)
# crash %>% mutate(POSTED_SPEED_LIMIT == recode(POSTED_SPEED_LIMIT, 
#                                               1:7 = 5,
#                                               8:12 = 10)) 
#                  POSTED_SPEED_LIMIT == 13:17, 15, POSTED_SPEED_LIMIT | 
#                  POSTED_SPEED_LIMIT == 18:22, 20, POSTED_SPEED_LIMIT) | 
#                  POSTED_SPEED_LIMIT == 23:27 -> 25,
#                  POSTED_SPEED_LIMIT == 28:32 -> 30,
#                  POSTED_SPEED_LIMIT == 33:37 -> 35,
#                  POSTED_SPEED_LIMIT == 38:42 -> 40,
#                  POSTED_SPEED_LIMIT == 43:47 -> 45,
#                  POSTED_SPEED_LIMIT == 48:52 -> 50,
#                  POSTED_SPEED_LIMIT == 53:57 -> 55,
#                  POSTED_SPEED_LIMIT == 58:62 -> 60,
#                  POSTED_SPEED_LIMIT == 63:70 -> 65)
```

```{r}
#crashes per month
count <- crash %>% 
  group_by(year= year(CRASH_DATE), month= month(CRASH_DATE)) %>% 
  summarize(n=n())

count %>% 
  ggplot(aes(x = year, y = n)) +
  geom_bar(stat = "identity") +
  scale_x_continuous(name = "Number of Crashes Per Year")

```

```{r}
count$date <- mdy(paste0(count$month,"-1-",count$year)) #Date Column
count <- count %>% 
  rename(Date=date, Count=n, Month=month, Year=year)
hc <- highchart(type="stock") %>%
  hc_add_series(data = count,
                   type = "line", hcaes(x = Date,
                   y = Count, 
                   )) %>% 
    hc_xAxis(title = list(text="Date")) %>% 
  hc_colors("#007fff") %>% 
  hc_plotOptions(series=list(backgroundColor = "#535353")) %>% 
    hc_tooltip(borderColor = "black",
             pointFormat = "<b> Crashes</b>: {point.Count}<br>")
hc
```

Speed limits were not recorded in common increments of 5pmh. 
```{r}
# original crash speed limit data
crash %>% 
  group_by(speed_limit=POSTED_SPEED_LIMIT) %>% 
  summarize(n=n()) %>%
  kbl(caption = "Crashes Occuring In Each Speed Limit within Chicago") %>% 
  kable_classic( full_width = T) %>%
  column_spec(1,  # column 1
              bold=T,  # make it bold
              border_right = T) %>% # add the line on the right side of the column
  column_spec(2)
```

Now include only increments of 5mph:
```{r}
slc <- crash %>% 
  filter(POSTED_SPEED_LIMIT %in% seq(from=5, to=70, by=5)) %>% 
    group_by(speed_limit=POSTED_SPEED_LIMIT) %>% 
    summarize(n=n())
kbl(slc) %>%
  kable_classic(full_width = F) %>%
  add_header_above(c("MPH" = 1, "Crashes" = 1)) # number of items in each group and gives them a label

# cooler if you are trying to group multiple variables in a chart.
```
Really good link for Kable charts:
https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html

## Crashes within each posted speed limit
```{r}
n1 <- slc %>% 
  ggplot(aes(x = speed_limit, y = n)) +
  geom_bar(stat = "identity", fill="#68ccff")+
  scale_x_continuous(breaks = seq(5,70,5), labels = seq(5,70,5)) +
  geom_text(aes( label = n), vjust = -0.3) + 
  ylab("Crashes") +
  xlab("Posted Speed Limit (mph)") +
  theme(panel.background = element_rect(fill = "#e8f7ff"), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
n1
```

Bring in the Chicago Beats Shapefile for mapping!
```{r}
shapefile <- st_read("C:/Users/aleaw/OneDrive/Desktop/LessonPlans/LessonPlans_PPOL105/crashes-COVID19_files/geo_export_19c4ac07-2232-427b-9ba0-127c6c22cd08.shp")

```


# Map of Beat 3100 Highlighted
```{r}
beatcount <- crash %>% 
  group_by() %>% 
  summarize(count=n()) %>% 
  arrange(beat_num) # Creating a data set with the count of crashes in each beat

# practice map
tm_shape(shapefile) +
  tm_borders() +
  tm_shape(shapefile) +
  tm_fill(col = "blue", alpha=0.65)

beatmap <- merge(shapefile, crash)
```


# Details

- Who did you collaborate with: xxxx
- Approximately how much time did you spend on this problem set: xxxx
- What, if anything, gave you the most trouble: xxxx


# Load and wrangle data

Load two CSV files and join them to create a data frame named `crashes` (i.e. using `crashes` as the main data frame, left join `COVID19` to it)

```{r load-libraries-data}

#crashes <- read_csv("data/  .csv")
#COVID19 <- read_csv("data   .csv")
```

```{r wrangle-data}
# Do something to add the names of the religions
#religion_utah <- religions
```


# Answer questions

Answer the following questions. I occasionally provide a chunk of code to help you get started. 

## General religious landscape

### Top religions

What are the top 10 religions in Utah, based on number of adherents?

<!-- ```{r top-10-adherents} -->
<!-- top_10_adherents <- religion_utah %>%  -->
<!--   group_by(religion_clean) %>%  -->
<!--   summarize(total = sum(adherents, na.rm = TRUE)) %>%  -->
<!--   top_n(10, total) %>%  -->
<!--   arrange(desc(total)) -->

<!-- top_10_adherents -->
<!-- ``` -->

<!-- Make a bar plot of the top 10 religions in Utah, by adherents: -->

<!-- ```{r top-10-adherents-plot} -->
<!-- # Here we make a copy of the top_10_adherents data from above and we -->
<!-- # change/mutate the `religion_clean` column so that the religion labels are in -->
<!-- # order (that's what `fct_inorder()` does---it converts `religion_clean` into a -->
<!-- # list of categories and puts them in a specific order) -->
<!-- top_10_adherents_plot <- top_10_adherents %>%  -->
<!--   mutate(religion_clean = fct_inorder(religion_clean)) -->

<!-- # Notice the `fct_rev()` here. That takes the list of categories in -->
<!-- # `religion_clean` and reverses them. If they're not reversed, they'll be -->
<!-- # plotted from lowest to highest, and LDS will be at the bottom. -->
<!-- # -->
<!-- # Also notice the `coord_flip()` layer. This rotates the plot so the bars go -->
<!-- # horizontally intead of vertically -->
<!-- ggplot(top_10_adherents_plot, aes(x = fct_rev(religion_clean), y = total)) + -->
<!--   geom_col() +  -->
<!--   labs(x = "Denomination", y = "Number of adherents") + -->
<!--   coord_flip() -->
<!-- ``` -->


<!-- Which 10 religions have the most number of congregations? -->

<!-- ```{r top-10-congregations} -->

<!-- ``` -->

<!-- Make a bar plot of the top 10 religions in Utah, by congregations: -->

<!-- ```{r top-10-congregations-plot} -->

<!-- ``` -->


<!-- Why are Catholics ranked differently in the two tables? xxxx -->


<!-- ### Proportion of membership in total population -->

<!-- Utah's total population in 2010 was 2,763,885. Calculate the proportion of each of the top 10 religions based on this population (i.e. Catholics made up X percent of Utah's population in 2010). -->

<!-- Hint: copy the code from the `top-10-adherents` chunk above and use `mutate()` to create a new column. -->

<!-- ```{r top-10-adherents-percent} -->

<!-- ``` -->


<!-- ## LDS membership -->

<!-- Which 5 counties in Utah have the most LDS members (or adherents)? -->

<!-- ```{r top-5-lds-adherents} -->
<!-- religion_utah %>%  -->
<!--   filter(religion_clean == "LDS") %>%  -->
<!--   top_n(5, adherents) %>%  -->
<!--   arrange(desc(adherents)) %>%  -->
<!--   select(county, religion_clean, adherents) -->
<!-- ``` -->

<!-- Which 5 counties in Utah have the most LDS congregations? -->

<!-- ```{r top-5-lds-congregations} -->

<!-- ``` -->

<!-- Which 5 counties in Utah have the highest proportion of LDS members? (i.e. `adherents_per_1000`) -->

<!-- ```{r top-5-lds-proportion} -->

<!-- ``` -->

<!-- Why are these answers not all the same? xxxx -->


<!-- ## Non-LDS membership -->

<!-- What are the top 5 non-LDS religions in Utah, based on number of adherents? -->

<!-- ```{r top-5-non-lds-adherents} -->

<!-- ``` -->


<!-- Which 10 counties have the highest proportion (`adherents_per_1000`) of non-LDS religions? What are those religions? -->

<!-- ```{r top-10-non-lds-counties} -->

<!-- ``` -->

<!-- Which 5 counties have the highest proportion of Muslims? -->

<!-- ```{r top-5-muslims} -->

<!-- ``` -->

<!-- Which 5 counties have the highest proportion of Buddhists? -->

<!-- ```{r top-5-buddhists} -->

<!-- ``` -->