---
title: "Intro to R"
author: "Alea Wilbur"
date: "8/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
Package - collection of R functions, data and compiled code  

Library -The location where the packages are stored  

If  there is  a  particular  functionality  that  you  require,  you  can  download  the  package  from the appropriate  site,  and  it  will  be  stored  in  your  library.  To  use  the  package,use  the  command **library()** to  load  the  package  in  the  current  R  session.  Then  just  call  the  appropriate  package functions

_install.packages("package_name")_ – Install the package from CRAN repository

_install.packages(c("package_1", “"package_2", "package_3"))_ – Install multiple packages 

_library("package_name")_ – Load the package in current R session

# Intro to R - Objects, Vectors, etc.

First, R can function just like a calculator
```{r}
2+2
2-2
2*2
2/2
3%%2
```



R is an open source software whose packages are developed by several individuals around the world. 

There are some coordinated efforts (e.g., tidyverse) but, in general, distributed development means that uniform conventions are often not followed concerning function names, arguments, and documentation.

This means that there are several ways to "code" in R and get to the same output. 

## Objects

**Objects** hold information - numbers, text, images... Each object has a name and we can assign content to an object using **<-**. You can also use **=** but the arrow is generally preferred


Let's create an object storing the number 2

```{r }

object1 <- 2

#Now, let's create another object storing the number 3

object2 <- 3

```

See what happens when we sum them

```{r, tidy=TRUE}

object1+object2

# tidy = TRUE in the markdown file adds spaces to make the code "tidy" and easy to read
```

## Vectors

An type of object that stores multiple pieces of information is called **vector**.

**c** is the function that we use to **combine** multiple values into one object.

*A function is a command that take an object and perform an operation*


We can create an object called ob1 storing the following values: 1, 3, 4, 5, 5

```{r }
ob1 <- c(1, 3, 4, 5, 5)
ob1        # To inspect an object, you can just type its name.
```
I strongly encourage you to always inspect your objects, vectors, matrices... Checking on your objects helps catch mistakes early.



### Object names

- Object can have any name but you *cannot* use spaces in an object name (e.g., "Object A" -> "ObjectA")

- You can use a dot or an underscore to separate words (underscore is preferred)

- R is case sensitive

- **Always** give your objects a "good name"  
    - Intuitive / meaningful
    - Concise / short    
    - Easy to remember
    - Unique (e.g., do not name an object as a variable or another object)

**Good names**  
- data_1  
- data_2   
- crime_data  

**OK names**  
- data.1 (dot)  
- mydata (generic)  

**Bad names**  
- myfirstdata_nov222020 (too long)  
- my_very_first_ObjectInR (too complex)  

**Very bad names**   
- dsafdsafadf_daada (meaningless) 

## Strings

We can store multiple type of values into a vector, not just numbers.

**Strings** (or characters) are pieces of text information which are stored in quotation marks **""** 

```{r }

week_days <- c("monday", "tuesday", "wednesday", "thursday", "friday", "saturday", "sunday")

week_days

```

## Class 

To know the type of value stored into an object, we can use the command **class**. 

```{r }
class(week_days)
```

### Your turn! - Change example a little

Create two objects, called "names" and "yob".   
- names should be a string containing the following names: Mary, Mark, John, David, and Claudia  
- yob should be a numeric vector containing the following years of birth: 1990, 1987, 1980, 1985, 1993  

_Which is the class of each vector?_


#### Solution

```{r}

names <- c("Mary", "Mark", "John", "David", "Claudia")

yob <- c(1999, 1985, 1960, 1955, 1974)


class(names)

class(yob)

```

We now have two **vectors** containing multiple values. One vector is **character** and the other is **numeric**. 

> Make sure to leave a space after each comma


## Matrices

Two or more vectors can be combined into a matrix. 

In this case, we use either **rbind** or **cbind** to bind our data by row or column, respectively.
```{r}
cbind(names, yob)

rbind(names, yob)
```

*What is the difference between the two commands?*

Let's use cbind to create a matrix to 'play' with. We call it data_1.

```{r }

data_1 <- cbind(names, yob)

class(data_1)

```


### Matrix dimensions

Matrices are made of rows and columns. You can check how many rows or columns at any time using the following commands:

```{r } 

nrow(data_1)

ncol(data_1)

dim(data_1)


```

>**Important**: R always stores information in the **row-column** format

These commands should be the first one you use when opening a dataset in R - check if it is loaded correctly + data size should match your expectations (e.g., if a dataset is at the state level, it should have 50 observations)



### Matrix positions

Like in battleship, we can identify information contained in a matrix by their row-column position

```{r }

# Position: row 2, column 2
data_1[2, 2]

# Position: row 5, column 1
data_1[5, 1]

```

_What happens if you omit the column or the row number? E.g., data_1[2, ]?_


## Numerical indexing

```{r }

# No column is specified
data_1[2, ]

```

```{r }

# No row is specified
data_1[, 2]

```

We can use [] to subset an object

```{r }

data_1[, 2]

```

We can save the new subset in a new object

```{r }

data_1_subset = data_1[, 2]

data_1_subset

```

## Dataframes

We will rarely work with matrices. In most cases, we will use dataframes. A **dataframe** is a dataset in R. We can convert any object into a dataframe

```{r }

data_df <- as.data.frame(data_1)

class(data_df) 

class(data_1)

data_df

```
Note that we can use the same commands as before to check the dimension of the dataframe

```{r }

ncol(data_df)

nrow(data_df)

dim(data_df)

```

Dataframes have special properties. For instance, they have column and row names

```{r }

colnames(data_df)

rownames(data_df)

```


```{r}
# Data Frame
Numbers <- c(1, 2, 3, 4, 5) 
Alphabets <- c("A", "B", "C", "D", "E") 
Boolean <- c(TRUE, FALSE, TRUE, TRUE, FALSE)
Float <- c(1.1, 2.2, 3.3, 4.4, 5.5)
df <- data.frame(Numbers, Alphabets, Boolean, Float)
df
```

```{r}
# Analyzing a DataFrame

dim(df)
ncol(df)
nrow(df)
str(df)
names(df)
colnames(df)
rownames(df)

head(df,2)
tail(df,2)

```


### Columns

Since columns have names, we can call each column using the symbol **$**

```{r }

data_df$names

data_df$yob

```

Important: You always need to call both the dataset and the column name

*datasetName$ColumnName*

We can perform any operation on columns. Let's try with checking the class of those columns.


```{r }

class(data_df$names)

class(data_df$yob)

```


## What is a factor?

Special class of vectors for categorical variables. Factors are composed by levels (a.k.a., categories). R uses factors to represent categorical variables that have a known set of possible values.

- **factor:** dog, cat, cat, dog, bird, dog
- **levels:** dog, cat, bird

```{r }
data_df$yob
```


## Converting class

You can always convert any vector (or column or row) from one class to the other 

- **as.**numeric
- **as.**character
- **as.**factor
- **as.**matrix
- **as.**data.frame


Note: When you recode a variable, it's good practice to save it as a new one. That way if you make a mistake, the original data still exists.  

- It allows you to check your work 

- You might need to go back to the original variable

- If you make a mistake, you don't have to upload your dataset again

- You can always clean your dataframe at the end (e.g., keep only relevant columns)

### From factor to numeric

By using as.numeric, the new vector stores the # of the level but not their content. 

- **factor:** dog, cat, cat, dog, bird, dog
- **levels:** dog, cat, bird
- **level numbers:** 1, 2, 3

```{r }

data_df$yob

data_df$yob_R

```
We need to convert into numeric each level of the vector

```{r }

data_df$yob_R2 <- as.numeric(levels(data_df$yob))[data_df$yob]

class(data_df$yob_R2)

data_df$yob_R2

```

### Summary of Class types

**Class**    | **Description**                          | 
-------------| -----------------------------------------|
character    |It stores text information. 
numeric      |It stores numbers (continuous variables)
factors      |It stores categorical variables
levels       |It stores each category of a factor



## Operations with column

You can manipulate columns in the same way you would with vectors (mostly). 

For instance, we can create a new column called **age** where we calculate the age for each individuals in the current year.

```{r }

data_df$age <- 2021 - data_df$yob_R2

data_df$age

```

You can also decide to calculate the age in terms of months instead of years

```{r }

data_df_agemonths <- data_df$age * 12

```

In sum, you can easily perform operations with your columns.

## Functions

When learning about a new function, you generally want to retrieve three pieces of information:

- **Description** wthat the function does

- **Usage** how you are expected to write the function

- **Arguments** what each part of the function does.

All help pages also contain an **"examples"** section where you can see how the function is used in practice.

Even when you discover new functions from other sources, you should check out the help page to understand all possible options provided by the arguments. 

Let's use some descriptive statistics functions to check out the variable *age*. 

```{r, eval = F}

table(data_df$age) # Frequencies

mean(data_df$age) # Mean value

min(data_df$age) # Minimum value

max(data_df$age) # Maximum value

sd(data_df$age) # Standard deviation

median(data_df$age) # Median value

quantile(data, 0.25) # Quantile

```

## Logical indexing

It is possible that you might want to know the average income of only those individuals who were born before 1980. 

We can use [] to subset our dataframe. 

We are asking R to subset *data_df* and keep only those rows that have a value lower than 1980 in the column *yob_R2*. We are keeping all columns in the dataset as we are not specifying any columns after the comma.

```{r }

data_df[data_df$yob_R2 < 1980, ]

data_df2 <- data_df[data_df$yob_R2 < 1980, ] # We could save it as a separate dataframe


data_df2
```
We can use logical indexing to perform operations on a subset of observations. Try to calculate the mean age for those individuals born before 1980 (start with code to calculate the mean age for everyone and modify it).


```{r }
mean(data_df$age[data_df$yob_R2 < 1980])

mean(data_df$age[data_df$yob_R2 <= 1980])
```

Now calculate the mean age for those called "Mary". Note that equal is represented by the symbol **==** when used for logical indexing.

```{r}
mean(data_df$age[data_df$names == "Mary"])
```

### Recap

Terminology used in this class provides you the basics to talk about R concepts and elements (vectors, objects, functions...)

'dollar sign' syntax is so called because of the use of **$** to connect a dataframe name with a column name.

Dataframes are a very common way to work with data in R. Some functions do not work with tibbles (tidyverse database format) so you'll likely go back to this at one point (e.g., regression analysis classes)

Tidyverse is better for data wrangling and visualization. 

# UNIVARIATE ANALYSIS

```{r}

# Factor

library(datasets)
data("mtcars")
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$vs <- as.factor(mtcars$vs)
mtcars$am <- as.factor(mtcars$am)
mtcars$gear <- as.factor(mtcars$gear)
mtcars$carb <- as.factor(mtcars$carb)
str(mtcars)

tab <- table(mtcars$am)
tab

prop.table(tab)

summary(mtcars[,c("cyl" , "vs" , "am" , "gear" , "carb")])

```


```{r}
# Numeric

mean(mtcars$mpg)
median(mtcars$mpg)
min(mtcars$mpg)
max(mtcars$mpg)
range(mtcars$mpg)
quantile(mtcars$mpg)
summary(mtcars$mpg)

```

# BIVARIATE ANALYSIS

```{r}
# Two Factors
data(esoph)
View(esoph)
str(esoph)

tab <- table(esoph$agegp, esoph$alcgp)
tab

prop_tab <- prop.table(tab)
prop_tab

addmargins(tab)
addmargins(prop_tab)
round(prop_tab,2)
round(addmargins(prop_tab),2)

prop_tab1 <- prop.table(tab, margin=1)
round(addmargins(prop_tab1),2)

prop_tab2 <- prop.table(tab, margin=2)
round(addmargins(prop_tab2),2)
```


```{r}
# Two Numeric
data(mtcars)
View(mtcars)

cor(mtcars$mpg , mtcars$hp)
cor(mtcars$disp , mtcars$hp)
```


```{r}
# One Factor & One Numeric
aggregate(mpg ~ cyl, data = mtcars, FUN = "mean")
aggregate(mpg ~ cyl, data = mtcars, FUN = "sd")

mtcars %>% group_by(cyl) %>% summarise(avg=mean(mpg) , median=median(mpg) , std=sd(mpg))

describeBy(mtcars$mpg , mtcars$cyl)

```

# Probability
```{r}
#Random normal distribution
rnorm(mean=1000, sd=250, 100)

# Histogram
values <- rnorm(mean=1000, sd=250, 10000)
hist(values, main="Normal Distribution", col="steelblue", xlab="Mean=1000, SD=250", freq=F)
lines(density(values), col="yellow", lwd=3)

# Percentile given observed value
pnorm(65 , 50 , 10)

# Value given a percentile
qnorm(0.9 , 50 , 10)

```

